Triptych-2: efficient proofs for confidential transactions

### 2020-03-12 paper-multi/iacr.tex
Updated Triptych-2 paper

\title{Triptych-2: efficient confidential transactions}
\author{Sarang Noether \\ Monero Research Lab \\ \texttt{sarang.noether@protonmail.com}}
\date{\today}

@@

\begin{abstract}
Confidential transactions are used in distributed digital assets to demonstrate the balance of values hidden in commitments, while retaining signer ambiguity.
In a previous paper, we describe how to use signer-ambiguous proofs of knowledge of the opening of multiple commitments to zero; these proofs can be combined to build confidential transactions, but require separate proofs when spending multiple input assets.
In this work, we extend our previous work to build Triptych-2, a proving system that uses a single proof to demonstrate transaction balance and the signing of multiple signer-ambiguous inputs at once.
While soundness depends on a novel dual discrete-logarithm hardness assumption, we use data from the Monero blockchain to show that Triptych-2 provides faster total batch verification time than other trust-free confidential transaction constructions.
\end{abstract}

\section{Introduction}
Distributed digital assets, starting with Bitcoin, rely on digital signatures to authorize transactions and transfer the spend authority of funds.
While early protocols based on the model established by Bitcoin have the advantage of simplicity, they lack desirable properties relating to privacy and indistinguishability.
In particular, the address/signature graph formed by the blockchain in such assets is trivially traceable, and the amounts a signature authorizes in a transaction are visible.

To protect users of distributed assets and permit better indistinguishability of transactions for fungibility, several methods have been proposed and deployed toward confidential transactions.
The CryptoNote protocol, an early proposal for obscuring transaction graphs, relied on linkable ring signatures on transaction outputs with denominated amounts \cite{cryptonote}
A later advancement to the RingCT protocol used Pedersen commitments to amounts associated with transaction outputs and extended a ring signature construction by Liu \text{et al.} to support parallel signatures in a matrix arrangement \cite{ringct,liu}.

@@

Until recently, transaction protocols desiring robust signer ambiguity and flexible direct anonymous transfers of arbitrary amounts faced a tradeoff between a trust-free construction and efficiency in proof sizes and verification complexity.
Recent work in this area relies on one of two novel proving systems to build transactions that scale in size logarithmically with the size of the prover-specified anonymity set, while requiring no trusted setup.
A one-of-many commitment-to-zero proving system of Groth and Kohlweiss was used to build simple ring signatures and Zerocoin-style transactions, and was later extended to support accountable ring signatures more efficiently \cite{groth,bootle}.
This proving system forms the basis of Lelantus \cite{lelantus}, which extends the Zerocoin-style transaction model of \cite{groth} by using multiple-base Pedersen commitments to incorporate amounts alongside serial numbers; a modification to the proofs shows balance, and multiple proofs are needed for typical transactions.
However, Lelantus transactions allow a transaction sender to later identify when the recipient spends their funds; a mitigation to this solves the tracing problem, but at the cost of eliminating a useful one-time addressing construction that ensures recipient anonymity.
The same proving system in \cite{groth} is used in Triptych \cite{triptych}, a multi-dimensional linkable ring signature construction that can be used to build confidential transactions similarly to \cite{ringct}; Triptych notably supports one-time addressing and arbitrary amounts, while still requiring the use of multiple proofs for transactions spending multiple transaction outputs.
Interestingly, recent independent work extends \cite{groth} to support proofs on multiple commitments in a single list (as we also introduce below), with a particular application to Zether \cite{zether}; however, this uses a different approach than we take here, and results in larger proofs.

The Bulletproofs \cite{bulletproofs} proving system uses an inner-product compression method to build range and circuit satisfiability proofs with logarithmic size scaling.
Besides the increasingly widespread deployment for commitment range proofs, the underlying structure of Bulletproofs is used to build proving systems for confidential transactions.
Omniring \cite{omniring} uses similar methods to build transactions that demonstrate spend authorization for multiple signer-ambiguous transaction inputs within a single proof, while directly integrating range proving.
The result is very small proofs; however, verification is slower than other approaches, and requirements for unique group generators mean such proofs cannot be verified in efficient batches.
RingCT 3.0 \cite{rct3} also uses a Bulletproof-style construction.
In an early version, transactions required separate proofs for multiple inputs; however, an update incorporates a single proof for all inputs to a transaction.
Notably, this construction admits efficient batch verification of separate proofs; however, a limitation on the prover requires that transactions use a power-of-2 number of inputs that yields extremely poor scaling for larger transactions.

\subsection{Our contribution}
We extend the proving system in Triptych \cite{triptych} in two significant ways, which we call Triptych-2.
First, we enable the prover to show knowledge of multiple signing keys in a single proof simultaneously using one set of combined proof elements; unlike \cite{rct3}, our modification admits any number of transaction inputs without limitations.
Second, we show Pedersen commitment balance directly within the same proof; in particular, we show that a particular combination of input and output commitments sums to a commitment to zero.
We note that the soundness of the construction depends on a novel dual discrete-logarithm hardness assumption that we are not able to reduce to a standard hardness assumption; while we consider this novel assumption reasonable, it is untested.

Taken together, these changes produce a useful transaction protocol intended for efficient use in blockchain confidential transactions.
A transaction can sign for multiple signer-ambiguous inputs and prove that the transaction balances using a single proof, a major improvement to currently-deployed constructions that require multiple proofs and separate auxiliary Pedersen amount commitments.
Further, we use actual blockchain data from the Monero digital asset project to directly compare the size and verification time of Triptych-2 to that of Triptych and two separate RingCT 3.0 constructions from earlier work \cite{triptych,rct3,lelantus}.
We find that Triptych-2 provides superior verification performance compared to other trust-free confidential transaction constructions.

\section{Preliminaries}
\subsection{Public parameters}
Let $\G$ be a cyclic group in which the discrete logarithm problem is hard, and let $\F$ be its scalar field.
Let $\hs: \{0,1\}^* \to \F$ and $\hp: \{0,1\}^* \to \G$ be cryptographic hash functions.
Assume that any point of the form $G_i$ or $H_i$ (possibly with multiple indices) is a generator of $\G$ whose discrete logarithm relationship with all other such points is unknown.
Let $G, H$ be distinguished generators of $\G$ used for Pedersen commitments.
Let $U$ be a distinguished generator of $\G$.
Let $N = n^m$ be a positive integer.
Note that all such generators may be produced using public randomness; for example, the use of a suitable hash function with domain separation may be appropriate.
All such public parameters are assumed to comprise a global reference string known to all players; in particular, we generally exclude explicit reference to public parameters from algorithm definitions and Fiat-Shamir transcript hashes for readability.

\subsection{Tensor commitment}
Let $\com$ be an additively homomorphic commitment scheme that is perfectly hiding and (at least) computationally binding.
In this work, we assume use of a simple extension to the Pedersen commitment scheme: for $x,r \in \F$, define $\com(x,r) \equiv xG + rH$ to be the commitment of the value $x$ with randomness $r$.

For a three-dimensional tensor $f \equiv (f_{i,j,k}) \subset \F$ and blinding factor $r \in \F$, define the Pedersen tensor commitment
$$\com(f,r) \equiv rH + \sum_{i,j,k} f_{i,j,k} G_{i,j,k}$$
using fixed independent generators as described above.

@@

We sometimes use index subscript notation of the form $i_j$ to indicate the $j$ digit of $i$, where such a decomposition of $i$ is taken base $n$ with padded length $m$:
$$\sum_{j=0}^m i_j n^j = i$$
This notation will be specified explicitly where confusion may occur.

\subsection{Sigma protocols}
For a given relation $\mathcal{R}$, a sigma protocol for $\mathcal{R}$ is an interactive challenge-response protocol between a prover and verifier, where the prover wishes to convince the verifier of the validity of an instance of the relation.

A sigma protocol is complete, sound, and zero-knowledge.
These definitions are well known and found, for example, in \cite{groth}.
Essentially, these properties are:
\begin{itemize}
\item \textit{Perfectly complete}: For a witness to a statement in $\mathcal{R}$, an honest prover always convinces an honest verifier of the validity.
\item \textit{Special sound}: Given a statement in $\mathcal{R}$, if the prover can produce valid proof transcripts in response to multiple verifier challenges, then it is possible to extract a witness for the statement from the transcripts.
\item \textit{Special honest-verifier zero knowledge}: Given a statement and verifier challenge, it is possible to produce a simulated transcript without knowledge of a corresponding valid witness.
\end{itemize}
Sigma protocols can be made non-interactive by replacing random verifier challenges with hash-based transcript challenges under the random oracle model; this approach is called the Fiat-Shamir heuristic \cite{fiat}.

@@

\begin{definition}[Dual-target discrete logarithm]
\label{def:dual}
Let $\G$ be a group where the discrete logarithm problem is hard, and let $\F$ be its scalar field.
Let $n > 0$.
Consider the following game between a challenger and a probabilistic polynomial-time player $\A$:
\begin{itemize}
\item The challenger chooses $G,H \in \G$ uniformly at random and sends both to $\A$.
\item $\A$ chooses and returns sets $\{G_i\}_{i=0}^{n-1},\{H_i\}_{i=0}^{n-1} \subset \G$ to the challenger.
\item The challenger chooses a set $\{\mu_i\}_{i=0}^{n-1} \subset \F$ uniformly at random and sends it to $\A$.
\item $\A$ chooses and returns a set $\{x_i\}_{i=0}^{n-1} \subset \F$ to the challenger.
\end{itemize}
We say that $\A$ wins the $(n,\epsilon,t)$-dual-target discrete logarithm game if, in time less than $t$ and with probability less than $\epsilon$, the following are true:
\begin{itemize}
\item $\sumi \mu_i \left( G_i - x_iG \right) = 0$
\item $\sumi \mu_i \left( H - x_iH_i \right) = 0$
\item There exists an index $0 \leq i < n$ such that either $x_iG \neq G_i$ or $x_iH_i \neq H$.
\end{itemize}
\end{definition}

\section{Proving system}
We now present a sigma protocol for the following relation:
\begin{multline*}
\mathcal{R} \equiv \Bigg\{ \{M_i\}_{i=0}^{N-1},\{P_i\}_{i=0}^{N-1},\{J^{(u)}\}_{u=0}^{w-1},\{Q_j\}_{j=0}^{T-1} \subset \G \: ; \: \left( \{l^{(u)}\}_{u=0}^{w-1}, \{r^{(u)}\}_{u=0}^{w-1}, y \right) : \\
M_{l^{(u)}} = r^{(u)}G \; \forall \: u \in [0,w) \text{ and } r^{(u)}J^{(u)} = U \; \forall \: u \in [0,w) \text{ and } \sum_{u=0}^{w-1} P_{l^{(u)}} - \sum_{j=0}^{T-1} Q_j = yG \Bigg\}
\end{multline*}
This captures the necessary elements for transaction authorization, which we describe later; knowledge of secret keys $\{r^{(u)}\}$ will show that the prover has signing authority for inputs to a transaction, and knowledge of the secret key $y$ to an amount commitment difference will show that the transaction balances.
Finally, comparison of the verifiable random function used to produce $\{J^{(u)}\}$ will be used to prevent attempts to sign for the same secret key without revealing the corresponding public key, which is important for signer ambiguity and double-spend protection within or between transactions.

Figures \ref{fig:sigma} and \ref{fig:sigma2} describes the prover and verifier interaction.
Next, we prove that the prover and verifer routines constitute a sigma protocol that is perfectly correct, special sound, and special honest-verifier zero knowledge.

@@

\begin{proof}
The proof proceeds similarly to that of \cite{triptych}.

To show the protocol is perfectly complete, suppose a verifier receives an honest proof.
Equation \ref{eqn:ab} holds from the identity $$\sumi \sigma_{j,i}^{(u)} = 1$$ for all $0 \leq j < m$ and $0 \leq u > w$.
Similarly, Equation \ref{eqn:cd} holds from the identity $$(\sigma_{j,i}^{(u)})^2 = \sigma_{j,i}^{(u)}$$ for all $0 \leq i < n$ and $0 \leq j < m$ and $0 \leq u < w$.

@@

We now show that the sigma protocol is $(m+1)$-special sound for $m > 1$.
To do so, we will produce an extractor that, when provided a set of $m+1$ valid responses to corresponding distinct $m+1$ verifier challenges for the same statements, produces a witness satisfying the relation $\mathcal{R}'$.

@@

\section{Transaction model}
We now describe how to apply the sigma protocol for $\mathcal{R}$ in a confidential transaction model.
In such a model, funds are represented by so-called \textit{outputs} generated in transactions, which become inputs for later transactions.
Notably, in a \textit{confidential} transaction model, we wish to introduce ambiguity as to which transaction inputs are being consumed, as well as the corresponding values represented by them.

Outputs are constructed as Pedersen commitments to zero, such that an output is of the form $M \equiv rG$ with signing key $r$.
We do not consider here the method by which output keys are generated, but assume this is offloaded to another part of the transaction model.
Each output comes equipped with a value commitment $P \equiv sG + aH$, where $a$ is the amount and $s$ a blinding factor.

A user of the protocol forms an input anonymity set $\{M_i\}_{i=0}^{N-1}$ of outputs from previous transactions.
Of these, suppose that the set $\{l^{(u)}\}_{u=0}^{w-1}$ represents the indices of the $w$ outputs for which the user wishes to sign; that is, the user knows $\{r^{(u)}\}_{u=0}^{w-1}$ such that $r^{(u)}G = M_{l^{(u)}}$ for all $0 \leq u < w$.
To help ensure ambiguity of the signing indices, assume the user has shuffled the input anonymity set.
Each element of the anonymity set has a corresponding value commitment $\{P_i \equiv s_iG + a_iH\}_{i=0}^{N-1}$.

Further, the user generates a set of new outputs $\{Q_j\}_{j=0}^{T-1}$ produced by the transaction, with corresponding value commitments of the form $\{t_jG + b_jH\}_{j=0}^{T-1}$.
Because the transaction value must balance, the following must hold: $$\sum_{u=0}^{w-1} a_{l^{(u)}} = \sum_{j=0}^{T-1} b_j$$

@@

We call the set $\{J^{(u)}\}_{u=0}^{w-1}$ the \textit{linking tags} for the transaction.
Similar to their use in linkable ring signatures, they are used by the verifier to detect attempts to sign with a secret key multiple times, either within the same transaction or between multiple transactions.
The construction $J^{(u)} = (r^{(u)})^{-1}U$ is an injective one-way verifiable random function \cite{dodis} using the secret key $r^{(u)}$; if the verifier sees the same linking tag used elsewhere, it knows that the (unknown) secret key was used in another proof.
In our confidential transaction model, this corresponds to an attempt to double-spend funds that would be rejected.

Since the sigma protocol is special honest-verifier zero knowledge, it is witness indistinguishable.
Note, however, that an observer who sees an input anonymity set containing a commitment for which the observer knows the opening can trivially test linking tags using its own secret keys to determine if they were used in the proof.
Therefore, witness indistinguishability is assumed to apply among input commitments for which an observer does not know a corresponding opening.

@@

We follow the same notation as before: $N$ is the size of the input anonymity set, $w$ is the number of signing indices, and $T$ is the number of output commitments used in the balance check.
Assuming that both elements of $\G$ and $\F$ occupy $32$ bytes of storage (as is the case for common elliptic curve groups using compressed point representations), we compare the size of proofs for RingCT 3.0 (two variants) \cite{rct3}, Triptych \cite{triptych}, and this work.
Note that we make some assumptions in order to produce fair comparisons here.
In Triptych and the original version of RingCT 3.0, certain commitment offsets are required as part of balance testing, which we consider, along with the multiple proofs needed in a transaction that wishes to sign for multiple inputs.
However, we ignore input set representation (which depends on implementation), range proofs (which are required for all protocols and not part of the proving systems), and any auxiliary data that would otherwise appear as part of transactions.
Table \ref{table:size} shows the results.

@@

Since each compared protocol scales differently with the transaction parameters, it is instructive to examine the overall size effects under real-world conditions.
To do this, we examine data from the Monero blockchain; in particular, for each transaction from 18 October 2018 to 14 February 2020, we extract the number of signed inputs, the number of newly-generated outputs, and the size of the anonymity set (in this analysis, it was constant).
Coinbase transactions, where new assets are generated by mining on a per-block basis, are ignored since they do not require this type of proof.
Figure \ref{fig:size} shows this data run against the values in \ref{table:size} with additional transaction auxiliary data (including range proofs and recipient-specific data) for increasing anonymity set size $N$.
Since all the protocols scale approximately linearly with $N$, overall verification time based on this parameter is an important bounding factor.

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{size.png}
\caption{Total chain growth by anonymity set size}
\label{fig:size}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{time.png}
\caption{Total verification time by anonymity set size}
\label{fig:time}
\end{figure}

While the updated RingCT 3.0 protocol from \cite{rct3} yields superior space scaling, its padding requirement that $w$ be a power of $2$ (padded to reach such a value) results in poor verification time scaling.
Comparatively, our construction results in lower overall verification complexity at the expense of overall size scaling.

>>>>

\title{Triptych-2: efficient proofs for confidential transactions}
\author{Sarang Noether \\ Monero Research Lab \\ \texttt{sarang.noether@protonmail.com}}
\date{\today}

@@

\begin{abstract}
Confidential transactions are used in distributed digital assets to demonstrate the balance of values hidden in commitments, while retaining signer ambiguity.
Previous work describes a signer-ambiguous proof of knowledge of the opening of commitments to zero at the same index across multiple public commitment sets and the evaluation of a verifiable random function used as a linking tag, and uses this to build a linkable ring signature called Triptych that can be used as a building block for a confidential transaction model.
In this work, we extend Triptych to build Triptych-2, a proving system that proves knowledge of openings of multiple commitments to zero within a single set, correct construction of a verifiable random function evaluated at each opening, and value balance across a separate list of commitments within a single proof.
While soundness depends on a novel dual discrete-logarithm hardness assumption, we use data from the Monero blockchain to show that Triptych-2 can be used in a confidential transaction model to provide faster total batch verification time than other state-of-the-art constructions without a trusted setup.
\end{abstract}

\section{Introduction}
Distributed digital assets, starting with Bitcoin, rely on digital signatures to authorize transactions and transfer the spend authority of funds.
While early protocols based on the model established by Bitcoin have the advantage of simplicity, they lack desirable properties relating to privacy and indistinguishability.
In particular, the address and signature graph formed by the blockchain in such assets is trivially traceable, and the amounts a signature authorizes in a transaction are visible.

To protect users of distributed assets and permit better indistinguishability of transactions for fungibility, several methods have been proposed and deployed toward confidential transactions.
The CryptoNote protocol, an early proposal for obscuring transaction graphs, relies on linkable ring signatures on transaction outputs with denominated amounts \cite{cryptonote}; a transaction includes a separate signature for each such amount, yielding poor scaling.
A later advancement to the RingCT protocol \cite{ringct} used Pedersen commitments to amounts associated with transaction outputs and extended a ring signature construction by Liu \text{et al.} \cite{liu} to support parallel signatures in a matrix arrangement, removing the need for denominated transaction outputs altogether.
However, both of these constructions use linkable ring signatures that scale linearly in size with the size of the anonymity set used in each signature.

@@

Until recently, transaction protocols desiring robust signer ambiguity and flexible direct anonymous transfers of arbitrary amounts faced a tradeoff between a trust-free construction and efficiency in proof sizes and verification complexity.
Recent work in this area generally relies on one of two novel proving systems to build transactions that scale in size logarithmically with the size of the prover-specified anonymity set, while requiring no trusted setup.
A one-of-many commitment-to-zero proving system of Groth and Kohlweiss \cite{groth} was used to build simple ring signatures and Zerocoin-style transactions, and was later extended to support accountable ring signatures more efficiently by Bootle \cite{bootle}.
This proving system forms the basis of Lelantus \cite{lelantus}, which extends the Zerocoin-style transaction model of \cite{groth} by using multiple-base Pedersen commitments to incorporate amounts alongside serial numbers; a modification to the proofs shows balance, and multiple proofs are needed for typical transactions.
However, Lelantus transactions allow a transaction sender to later identify when the recipient spends their funds; a mitigation to this solves the tracing problem, but at the cost of eliminating a useful one-time addressing construction that ensures recipient anonymity.
The same proving system in \cite{groth} is used in Triptych \cite{triptych}, a multi-dimensional linkable ring signature construction that can be used to build confidential transactions similarly to \cite{ringct}; Triptych notably supports one-time addressing and arbitrary amounts, while still requiring the use of multiple proofs for transactions spending multiple transaction outputs.
Interestingly, recent independent work extends \cite{groth} to support proofs on multiple commitments in a single list (as we also introduce below), with a particular application to Zether \cite{zether}; however, this uses a different approach than we take here, and seems to result in larger proofs.

The Bulletproofs \cite{bulletproofs} proving system uses an inner-product compression method to build range and circuit satisfiability proofs with logarithmic size scaling.
Besides an increasingly widespread deployment for commitment range proofs, the underlying structure of Bulletproofs is used to construct more specialized proving systems for confidential transactions.
Omniring \cite{omniring} uses this method to build transactions that demonstrate spend authorization for multiple signer-ambiguous transaction inputs within a single proof, while directly integrating range proving.
The result is very small proofs; however, verification is slower than other approaches, and requirements for unique group generators mean such proofs cannot be verified in efficient batches.
RingCT 3.0 \cite{rct3} also uses a Bulletproof-style construction.
In an early version, transactions required separate proofs for multiple inputs; while an update incorporates a single proof for all inputs to a transaction, it also requires that the number of inputs be padded to a power of two, which can yield poor verification complexity.

\subsection{Our contribution}
We extend the proving system in Triptych \cite{triptych} in two significant ways, which we call Triptych-2.
First, we enable the prover to show knowledge of multiple signing keys in a single proof simultaneously using one set of combined proof elements; unlike \cite{rct3}, our modification admits any number of transaction inputs without limitations.
We retain the verifiable random function evaluation that produces linking tags, which are required for the detection of repeat signing with the same commitment opening across proofs.
Second, we show Pedersen commitment value balance directly within the same proof; in particular, we show that a particular combination of input and output commitments sums to a commitment to zero.
We note that the soundness of the resulting proving system depends on a novel dual discrete-logarithm hardness assumption that we are not able to reduce to a standard hardness assumption; while we consider this novel assumption reasonable, it is untested.

Taken together, these changes have application to a transaction protocol for efficient use in blockchain confidential transactions that does not require any trusted setup process.
A transaction can sign for multiple signer-ambiguous inputs and prove that the transaction balances using a single proof, a major improvement to currently-deployed constructions that require multiple proofs and separate auxiliary Pedersen amount commitments.
Further, we use actual blockchain data from the Monero digital asset project to directly compare the overall size and verification time of Triptych-2 to that of Triptych and two separate RingCT 3.0 constructions.
We find that Triptych-2 provides superior verification performance compared to other trust-free confidential transaction constructions, with competitive size scaling.

\section{Preliminaries}
\subsection{Public parameters}
Let $\G$ be a cyclic group in which the discrete logarithm problem is hard, and let $\F$ be its scalar field.
Let $\hs: \{0,1\}^* \to \F$ and $\hp: \{0,1\}^* \to \G$ be cryptographic hash functions.
Let $N = n^m$ be a positive integer, where $m > 1$ (for our construction, we will use $n = 2$.
Assume that $G$, $H$, $U$, and any point of the form $G_i$ or $H_i$ (possibly with multiple indices) are uniformly random generators of $\G$ whose discrete logarithm relationship to each other is unknown.
Note that all such generators may be produced using public randomness; for example, the use of a suitable hash function with domain separation may be appropriate.
All such public parameters are assumed to comprise a global reference string known to all players; in particular, we generally exclude explicit reference to public parameters from algorithm definitions and Fiat-Shamir transcript hashes for readability.

\subsection{Tensor commitment}
Let $\com$ be an additively homomorphic commitment scheme that is perfectly hiding and (at least) computationally binding.
In this work, we assume use of a simple extension to the Pedersen commitment scheme: for $x,r \in \F$, define $\com(x,r) \equiv rG + xH$ to be the commitment of the value $x$ with randomness $r$.

For a three-dimensional tensor $f \equiv (f_{i,j,k}) \subset \F$ and blinding factor $r \in \F$, define the Pedersen tensor commitment
$$\com(f,r) \equiv rG + \sum_{i,j,k} f_{i,j,k} G_{i,j,k}$$
using fixed independent generators as described above.

@@

We sometimes use index subscript notation of the form $i_j$ to indicate the $j$ digit of $i$, where such a decomposition of $i$ is taken base $n$ with padded length $m$:
$$\sum_{j=0}^{m-1} i_j n^j = i$$
This notation will be specified explicitly where confusion may occur.

\subsection{Sigma protocols}
For a given relation $\mathcal{R}$, a sigma protocol for $\mathcal{R}$ is an interactive challenge-response protocol between a prover and verifier, where the prover wishes to convince the verifier it knows a witness corresponding to a statement in $\mathcal{R}$.

A sigma protocol is complete, sound, and zero-knowledge.
These definitions are well known and found, for example, in \cite{groth}.
Essentially, these properties are:
\begin{itemize}
\item \textit{Perfectly complete}: For a witness to a statement in $\mathcal{R}$, an honest prover always convinces an honest verifier of the validity.
\item \textit{Special sound}: Given a statement in $\mathcal{R}$, if the prover can produce valid proof transcripts in response to multiple verifier challenges, then it is possible to extract a witness for the statement from the transcripts.
\item \textit{Special honest-verifier zero knowledge}: Given a statement and known verifier challenge, it is possible to produce a simulated transcript without knowledge of a corresponding valid witness.
\end{itemize}
Sigma protocols can be made non-interactive by replacing random verifier challenges with hash-based transcript challenges under the random oracle model; this approach is called the Fiat-Shamir heuristic \cite{fiat}.

@@

\begin{definition}[Dual-target discrete logarithm problem]
\label{def:dual}
Let $\G$ be a group in which the discrete logarithm problem is hard, and let $\F$ be its scalar field.
Let $n > 0$.
Consider the following game between a challenger and a probabilistic polynomial-time player $\A$:
\begin{itemize}
\item The challenger chooses $G,H \in \G$ uniformly at random and sends both to $\A$.
\item $\A$ chooses and returns sets $\{G_i\}_{i=0}^{n-1},\{H_i\}_{i=0}^{n-1} \subset \G$ to the challenger.
\item The challenger chooses a set $\{\mu_i\}_{i=0}^{n-1} \subset \F$ uniformly at random and sends it to $\A$.
\item $\A$ chooses and returns a set $\{x_i\}_{i=0}^{n-1} \subset \F$ to the challenger.
\end{itemize}
We say that $\A$ wins the $(n,\epsilon,t)$-dual-target discrete logarithm game if, in time less than $t$ and with probability at least $\epsilon$, the following are true:
\begin{itemize}
\item $\sumi \mu_i \left( G_i - x_iG \right) = 0$
\item $\sumi \mu_i \left( H - x_iH_i \right) = 0$
\item There exists an index $0 \leq i < n$ such that either $x_iG \neq G_i$ or $x_iH_i \neq H$.
\end{itemize}
\end{definition}

\section{Proving system}
We now present a sigma protocol for the following relation:
\begin{multline*}
\mathcal{R} \equiv \Bigg\{ \{M_i\}_{i=0}^{N-1},\{P_i\}_{i=0}^{N-1},\{J^{(u)}\}_{u=0}^{w-1},\{Q_j\}_{j=0}^{T-1} \subset \G \: ; \: \left( \{l^{(u)}\}_{u=0}^{w-1}, \{r^{(u)}\}_{u=0}^{w-1}, y \right) : \\
M_{l^{(u)}} = r^{(u)}G \; \forall \: u \in [0,w) \text{ and } r^{(u)}J^{(u)} = U \; \forall \: u \in [0,w) \text{ and } \sum_{u=0}^{w-1} P_{l^{(u)}} - \sum_{j=0}^{T-1} Q_j = yG \Bigg\}
\end{multline*}
This captures the necessary elements for transaction authorization, which we describe later; knowledge of signing keys $\{r^{(u)}\}$ will show that the prover has signing authority for inputs to a transaction, and knowledge of the secret key $y$ to an amount commitment difference will show that the transaction amounts balance.
Finally, comparison of the verifiable random function used to produce $\{J^{(u)}\}$ will be used to prevent attempts to sign for the same secret key without revealing the corresponding public key, which is important for signer ambiguity and double-spend protection within or between proofs.

Figures \ref{fig:sigma} and \ref{fig:sigma2} describe the prover and verifier interaction.
Next, we prove that the prover and verifer routines constitute a sigma protocol that is perfectly correct, special sound, and special honest-verifier zero knowledge.

@@

\begin{proof}
The proof proceeds similarly to that of \cite{triptych}, which in turn follows the methods in \cite{groth,bootle}.

To show the protocol is perfectly complete, suppose a verifier receives an honest proof.
We wish to show that all verifier equations hold.
Equation \ref{eqn:ab} holds from the identity $$\sumi \sigma_{j,i}^{(u)} = 1$$ for all $0 \leq j < m$ and $0 \leq u > w$.
Similarly, Equation \ref{eqn:cd} holds from the identity $$(\sigma_{j,i}^{(u)})^2 = \sigma_{j,i}^{(u)}$$ for all $0 \leq i < n$ and $0 \leq j < m$ and $0 \leq u < w$.

@@

@@

\section{Transaction model}
We now describe how to apply the sigma protocol for $\mathcal{R}$ in a confidential transaction model.
In such a model, a transaction consumes so-called \textit{outputs} generated in previous transactions, and produces new outputs for use in later transactions.
Notably, in a \textit{confidential} transaction model, we wish to introduce ambiguity as to which transaction inputs are being consumed, as well as the corresponding values represented by them.

Outputs are constructed as Pedersen commitments to zero, such that an output public key is of the form $M \equiv rG$ with signing key $r$.
We do not consider here the method by which output keys are generated, but assume this is offloaded to another part of the transaction model.
Each output comes equipped with a value commitment $P \equiv sG + aH$, where $a$ is the amount and $s$ a blinding factor.

A user of the protocol forms an anonymity set $\{M_i\}_{i=0}^{N-1}$ of outputs from previous transactions.
Of these, suppose that the set $\{l^{(u)}\}_{u=0}^{w-1}$ represents the indices of the $w$ outputs for which the user wishes to sign; that is, the user knows $\{r^{(u)}\}_{u=0}^{w-1}$ such that $r^{(u)}G = M_{l^{(u)}}$ for all $0 \leq u < w$.
To help ensure ambiguity of the signing indices, assume the user has shuffled the anonymity set.
Each element of the anonymity set has a corresponding value commitment $\{P_i \equiv s_iG + a_iH\}_{i=0}^{N-1}$.

Further, the user generates a set of new outputs produced by the transaction, with corresponding value commitments of the form $\{Q_j \equiv t_jG + b_jH\}_{j=0}^{T-1}$.
Because the transaction value must balance, the following must hold: $$\sum_{u=0}^{w-1} a_{l^{(u)}} = \sum_{j=0}^{T-1} b_j$$
Note that there is no anonymity set for outputs generated by the transaction.

@@

We call the elements of the set $\{J^{(u)}\}_{u=0}^{w-1}$ the \textit{linking tags} for the transaction.
Similar to their use in linkable ring signatures, they are used by the verifier to detect attempts to sign with a secret key multiple times, either within the same transaction or between multiple transactions.
The construction $J^{(u)} = (r^{(u)})^{-1}U$ is an injective one-way verifiable random function \cite{dodis} using the secret key $r^{(u)}$; if the verifier sees the same linking tag used elsewhere, it knows that the (unknown) secret key was used again.
In our confidential transaction model, this corresponds to an attempt to double-spend funds that would be rejected by the verifier.

Since the sigma protocol is special honest-verifier zero knowledge, it is witness indistinguishable \cite{cramer}.
Note, however, that an observer who sees an input anonymity set containing a commitment for which the observer knows the opening can trivially test linking tags using its own secret keys to determine if they were used in the proof.
Therefore, witness indistinguishability is assumed to apply among input commitments for which an observer does not know a corresponding opening.

@@

We follow the same notation as before: $N = 2^m$ is the size of the input anonymity set for varying $m > 1$, $w$ is the number of signing indices, and $T$ is the number of output commitments used in the balance check.
Assuming that both elements of $\G$ and $\F$ occupy $32$ bytes of storage (as is the case for common elliptic curve groups using compressed point representations), we compare the size of proofs for RingCT 3.0 (two variants) \cite{rct3}, Triptych \cite{triptych}, and this work.
Note that we make some assumptions in order to produce fair comparisons here.
In Triptych and the original version of RingCT 3.0, certain commitment offsets are required as part of balance testing, which we consider, along with the multiple proofs needed in a transaction that wishes to sign for multiple inputs.
However, we ignore input set representation (which depends on implementation), range proofs (which are required for all protocols and not part of the proving systems), and any auxiliary data that would otherwise appear as part of transactions.
Table \ref{table:size} shows the results.

@@

Since each compared protocol scales differently with the transaction parameters, it is instructive to examine the overall size effects under real-world conditions.
To do this, we examine data from the Monero blockchain; in particular, for each transaction from 18 October 2018 to 14 February 2020, we extract the number of consumed outputs and the number of newly-generated outputs.
Coinbase transactions, where new assets are generated by mining on a per-block basis, are ignored since they do not require this type of proof.
Figure \ref{fig:size} shows this data run against the values in Table \ref{table:size} with additional transaction auxiliary data (including range proofs and recipient-specific data) for increasing anonymity set size $N$.
Since verification complexity in all the protocols scales approximately linearly with $N$, overall verification time based on this parameter is an important bounding factor.

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{size.png}
\caption{Total chain growth by anonymity set size}
\label{fig:size}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{time.png}
\caption{Total verification time by anonymity set size}
\label{fig:time}
\end{figure}

While the updated RingCT 3.0 protocol from \cite{rct3} yields superior space scaling, its requirement that $w$ be a power of two (padded to reach such a value) results in poor verification time scaling.
Comparatively, our construction results in lower overall verification complexity at the expense of overall size scaling.
